### 2021-02-22 ch02_scientific_background/scientific_background.tex
Finished chapter 06

The \emph{emergent constraint} approach tackles this problem by
\enquote{identifying robust, physically interpretable relationships between
  Earth system feedback behaviors on short, well-observed time scales and on
  time scales that span the twenty-first century and beyond}
\autocite{Eyring2019}. An illustration of the concept of emergent constraints
is shown in \cref{fig:02:emergent_constraint}. Each emergent constraint
requires two key components: an \emph{emergent relationship} and a
corresponding observation of the real world \autocite{Eyring2019}. The emergent
relationship (red line in \cref{fig:02:emergent_constraint}) is a robust and
physically interpretable inter-model relationship between a target variable $y$
related to the future climate and an observable $x$ of the past or present-day
climate. Basis for the relationship is output of the different climate models
of a multi-model ensemble (blue circles in \cref{fig:02:emergent_constraint}).
Using an observation of $x$, this emergent relationship can then be used to
derive a emergent constraint on $y$ (gray shaded area in
\cref{fig:02:emergent_constraint}) that considers uncertainties in the emergent
relationship itself (red shaded area in \cref{fig:02:emergent_constraint}) and
uncertainties in the observation (blue shaded area in
\cref{fig:02:emergent_constraint}).

\begin{figure}[t]
  \centering
  \includegraphics[width=\FigureWidth{}]{
    ch02_scientific_background/figs/emergent_constraint.pdf}
  \caption{Schematic illustration of the emergent constraint approach. Basis of
    every emergent constraint is a robust and physically interpretable emergent
    relationship (red line) between a target variable $y$ (\eg{} Earth system
    sensitivity or projection of future climate change) and an observable $x$
    (\eg{} past or present-day trend or variation) for the climate models of a
    multi-model ensemble (blue circles). With an observation of $x$,
    uncertainties in $y$ in the multi-model ensemble, illustrated by the yellow
    \acf{PDF}, can be reduced (gray \acs{PDF}). Uncertainties in the target
    variable $y$ arise from two sources: uncertainties in the observation (blue
    shaded area) and uncertainties in the emergent relationship (red shaded
    area). \SpringerAdaptedFrom{Eyring2019}.}
  \label{fig:02:emergent_constraint}
\end{figure}

>>>>

The \emph{emergent constraint} approach tackles this problem by
\enquote{identifying robust, physically interpretable relationships between
  Earth system feedback behaviors on short, well-observed time scales and on
  time scales that span the twenty-first century and beyond}
\autocite{Eyring2019}. An illustration of the concept of emergent constraints
is shown in \cref{fig:02:emergent_constraint}. Each emergent constraint
requires two key components: an \emph{emergent relationship} and a
corresponding observation of the real world \autocite{Eyring2019}. The emergent
relationship (red line in \cref{fig:02:emergent_constraint}) is a robust and
physically-interpretable inter-model relationship between a target variable $y$
related to the future climate and an observable $x$ of the past or present-day
climate. Basis for the relationship is output of the different climate models
of a multi-model ensemble (blue circles in \cref{fig:02:emergent_constraint}).
Using an observation of $x$, this emergent relationship can then be used to
derive a emergent constraint on $y$ (gray shaded area in
\cref{fig:02:emergent_constraint}) that considers uncertainties in the emergent
relationship itself (red shaded area in \cref{fig:02:emergent_constraint}) and
uncertainties in the observation (blue shaded area in
\cref{fig:02:emergent_constraint}).

\begin{figure}[t]
  \centering
  \includegraphics[width=\FigureWidth{}]{
    ch02_scientific_background/figs/emergent_constraint.pdf}
  \caption{Schematic illustration of the emergent constraint approach. Basis of
    every emergent constraint is a robust and physically-interpretable emergent
    relationship (red line) between a target variable $y$ (\eg{} Earth system
    sensitivity or projection of future climate change) and an observable $x$
    (\eg{} past or present-day trend or variation) for the climate models of a
    multi-model ensemble (blue circles). With an observation of $x$,
    uncertainties in $y$ in the multi-model ensemble, illustrated by the yellow
    \acf{PDF}, can be reduced (gray \acs{PDF}). Uncertainties in the target
    variable $y$ arise from two sources: uncertainties in the observation (blue
    shaded area) and uncertainties in the emergent relationship (red shaded
    area). \SpringerAdaptedFrom{Eyring2019}.}
  \label{fig:02:emergent_constraint}
\end{figure}

### 2021-02-22 ch05_paper_ecs/paper_ecs.tex
Finished chapter 06

As shown in the previous \namecrefs{sec:05:comparison_of_emergent_constraints},
most emergent relationships show smaller coefficients of determination when
evaluated on the new \acs{CMIP}6 ensemble compared to the \acs{CMIP}5 ensemble.
In this \namecref{sec:05:discussion}, we discuss possible reasons for, and
implications of, these differences. As reported by \textcite{Caldwell2014}, the
large amount of data provided by modern \acp{ESM} can generate spurious
correlations of variables between past climate and \ac{ECS} just by chance,
especially when only a small number of climate models is considered. This would
cause the performance of the emergent constraint to be reduced on out-of-sample
data (like the new \acs{CMIP}6 ensemble), since the emergent relationship
appeared just by chance and not because of a physically based mechanism.

>>>>

As shown in the previous \namecrefs{sec:05:comparison_of_emergent_constraints},
most emergent relationships show smaller coefficients of determination when
evaluated on the new \acs{CMIP}6 ensemble compared to the \acs{CMIP}5 ensemble.
In this \namecref{sec:05:discussion}, we discuss possible reasons for, and
implications of, these differences. As reported by \textcite{Caldwell2014}, the
large amount of data provided by modern \acp{ESM} can generate spurious
correlations of variables between past climate and \ac{ECS} just by chance,
especially when only a small number of climate models is considered. This would
cause the performance of the emergent constraint to be reduced on out-of-sample
data (like the new \acs{CMIP}6 ensemble), since the emergent relationship
appeared just by chance and not because of a physically-based mechanism.

### 2021-02-22 ch06_paper_gpp/paper_gpp.tex
Finished chapter 06

The spatial result of our \ac{GBRT} model in step 2a can be visualized by
comparing r\acs{MMM} and the output of the \acs{GBRT} model.
\Cref{fig:06:step2a_results:b} shows the bias of the rescaled \acs{CMIP}5
\acs{RCP}8.5 \ac{MMM} (r\acs{MMM}) compared to our \acs{GBRT} predicted
\ac{GPP} at the end of the \nth{21} century (\range{2091}{2100}), while
\Cref{fig:06:step2a_results:c} shows the bias of the historical \acs{CMIP}5
\ac{MMM} compared to the observational FLUXNET-MTE product \autocite{Jung2011}
averaged for the period \range{1991}{2000}. As reported by \textcite{Anav2013},
the historical \acs{CMIP}5 ensemble mean overestimates \ac{GPP} in most
regions, leading to a global mean of $138 \unit{\GtCyr{}}$, which is larger
than the $123 \unit{\GtCyr{}}$ estimate for the FLUXNET-MTE product. Regions
where \ac{GPP} is largely overestimated are the western parts of South America,
central and southern Africa, and East Asia. On the contrary, \ac{GPP} is
underestimated in small areas of Central America and northern parts of South
America. The bias patterns in
\cref{fig:06:step2a_results:b,fig:06:step2a_results:c} are very similar
(pattern correlation of $R^2 = 0.88$), which means that the \ac{GBRT} algorithm
detects regional biases in the historical simulation and corrects them in its
future predictions. This is also illustrated in
\cref{fig:06:step2a_results_biases:b}, in which many values are close to $1$,
while only a few negative values exist. However, we emphasize that our approach
is only to first order a bias correction (\ie{} subtracting the historical bias
from the \acs{RCP}8.5 \ac{MMM}). There are clear differences between
\cref{fig:06:step2a_results:b,fig:06:step2a_results:c}, which are also
illustrated in
\cref{fig:06:step2a_results_biases:a,fig:06:step2a_results_biases:b}. A simple
bias correction would only show constant values for the whole globe, whereas
our approach applies different corrections (in sign and magnitude) for
different regions. On the global domain, our approach predicts a \ac{GPP} at
the end of the century of $169 \unit{\GtCyr{}}$, which is consistent with the
global constraint from step 1 of $\pmrangeunit{171}{12}{\GtCyr{}}$. In summary,
our approach first corrects the \acs{CMIP}5 models response to \ac{CO2} (step
1) and second corrects the historical bias of the \acp{ESM} relative to
observations (step 2a).

@@

TBA.

>>>>

The spatial result of our \ac{GBRT} model in step 2a can be visualized by
comparing r\acs{MMM} and the output of the \acs{GBRT} model.
\Cref{fig:06:step2a_results:b} shows the bias of the rescaled \acs{CMIP}5
\acs{RCP}8.5 \ac{MMM} (r\acs{MMM}) compared to our \acs{GBRT}-predicted
\ac{GPP} at the end of the \nth{21} century (\range{2091}{2100}), while
\Cref{fig:06:step2a_results:c} shows the bias of the historical \acs{CMIP}5
\ac{MMM} compared to the observational FLUXNET-MTE product \autocite{Jung2011}
averaged for the period \range{1991}{2000}. As reported by \textcite{Anav2013},
the historical \acs{CMIP}5 ensemble mean overestimates \ac{GPP} in most
regions, leading to a global mean of $138 \unit{\GtCyr{}}$, which is larger
than the $123 \unit{\GtCyr{}}$ estimate for the FLUXNET-MTE product. Regions
where \ac{GPP} is largely overestimated are the western parts of South America,
central and southern Africa, and East Asia. On the contrary, \ac{GPP} is
underestimated in small areas of Central America and northern parts of South
America. The bias patterns in
\cref{fig:06:step2a_results:b,fig:06:step2a_results:c} are very similar
(pattern correlation of $R^2 = 0.88$), which means that the \ac{GBRT} algorithm
detects regional biases in the historical simulation and corrects them in its
future predictions. This is also illustrated in
\cref{fig:06:step2a_results_biases:b}, in which many values are close to $1$,
while only a few negative values exist. However, we emphasize that our approach
is only to first order a bias correction (\ie{} subtracting the historical bias
from the \acs{RCP}8.5 \ac{MMM}). There are clear differences between
\cref{fig:06:step2a_results:b,fig:06:step2a_results:c}, which are also
illustrated in
\cref{fig:06:step2a_results_biases:a,fig:06:step2a_results_biases:b}. A simple
bias correction would only show constant values for the whole globe, whereas
our approach applies different corrections (in sign and magnitude) for
different regions. On the global domain, our approach predicts a \ac{GPP} at
the end of the century of $169 \unit{\GtCyr{}}$, which is consistent with the
global constraint from step 1 of $\pmrangeunit{171}{12}{\GtCyr{}}$. In summary,
our approach first corrects the \acs{CMIP}5 models response to \ac{CO2} (step
1) and second corrects the historical bias of the \acp{ESM} relative to
observations (step 2a).

@@

In this paper, we develop a two-step approach to constrain the projected
\ac{GPP} at the end of the \nth{21} century (\range{2091}{2100}) in the
\acs{RCP}8.5 scenario. In the first step, we constrain the global mean \ac{GPP}
to $\pmrangeunit{171}{12}{\GtCyr{}}$ using a published emergent constraint
approach which we assume can also be applied to the global mean \ac{GPP} at the
end of the \nth{21} century in the \acs{RCP}8.5 scenario (see discussion in
\cref{subsec:06:step_1}). This step corresponds to the correction of the
\acp{ESM}' biases in the response of future \ac{GPP} to rising atmospheric
\ac{CO2} concentration, which is the main source of uncertainty in future
\ac{GPP} projections \autocite{Arora2013, Haverd2020, Rogers2017}. As already
discussed in \cref{subsec:06:step_1}, due to other than \ac{CO2} fertilization
drivers of the increase in the \ac{CO2} seasonal cycle amplitude
\autocite{Bastos2019, Forkel2016, Piao2018, Zhao2016} this emergent constraint
is not undisputed and can be replaced with another emergent constraint in
future studies. In the second step, a \ac{ML} approach is used to further
constrain the gridded \ac{GPP} based on process-based present-day predictors.
We consider two target variables: first (step 2a) the gridded monthly
climatologies of absolute \ac{GPP} (\range{2091}{2100}) and second (step 2b)
the gridded fractional \ac{GPP} change over the \nth{21} century (years
\range{2091}{2100} versus years \range{1991}{2000}; see \cref{eq:06:f}). Both
approaches give consistent results (see
\cref{fig:06:comparison_step2a_step2b}). The latter quantity shows an increased
\ac{GPP} change in the high latitudes compared to regions closer to the
equator, which might be attributed to an additional greening trend in the high
latitudes as supported by the feature map in
\cref{fig:06:step2b_ml_inference:c} that shows \ac{LAI} as dominant predictor
in this region. However, in the absence of a robust physical mechanism this
connection cannot be proven. For both target variables, the \ac{GBRT} algorithm
is superior to all other considered statistical models in terms of prediction
uncertainty evaluated in a leave-one-model-out \ac{CV} approach among different
statistical models. Compared to the unweighted \acs{CMIP}5 \ac{MMM} the mean
prediction error is reduced by $48 \unit{\%}$ (step 2a) and $16 \unit{\%}$
(step 2b); compared to a linear \ac{LASSO} regression model the mean prediction
error is reduced by approximately $3 \unit{\%}$ for both cases. However, local
standard errors are still large for both target variables, even for the
\ac{GBRT} model. Due to the unknown covariance structure a global aggregation
of these errors is not possible. Step 2a mainly corrects the bias of the
simulated absolute historical \ac{GPP} relative to observations
\autocite{Anav2013}. Consequently, the historical \ac{GPP} is by far the most
important predictor for future absolute \ac{GPP} (regions with high \ac{GPP} in
the past are likely to have high \ac{GPP} in the future and vice versa).
However, as we show in \cref{fig:06:step2a_results_biases}, the \ac{GBRT} model
expands this bias correction by taking more predictors than the historical
\ac{GPP} into account. A similar result is provided by step 2b:
\cref{fig:06:step2b_results_biases:b,fig:06:step2b_results_biases:c} show the
impact of the additional predictors as difference between the \acs{CMIP}5
\ac{MMM} and the \ac{GBRT} prediction of the \ac{GPP} change. In this case, the
target variable (fractional \ac{GPP} change) is already normalized with the
historical \ac{GPP}. Therefore, \ac{T} (second most important feature in step
2a) is now the most important predictor instead of historical \ac{GPP} (see
\cref{fig:06:step2b_ml_inference:d}). As shown in
\cref{fig:06:step2b_ml_inference:c}, the different features have also different
dominant regional impacts; but again, since there is no distinct physical
mechanism relating the predictors to the target variable this map requires
further analysis and needs to be treated with care.

Our \ac{GBRT} approach (step 2) is mathematically similar to \ac{MDER}
\autocite{Karpechko2013, Senftleben2020, Wenzel2016a}: We establish a
relationship between process-oriented, physically-relevant diagnostics and
future projections and then utilize this to predict today's observed conditions
into the future. We emphasize that the exact nature of this relationship is
strongly dependent on the climate model ensemble considered. Although derived
empirically, we argue that this relationship can still be used to extract
further information from the climate model ensemble for two reasons: First, we
consider process-oriented variables which are physically linked to \ac{GPP},
and second, we ensure its statistical robustness. The latter is achieved by
considering gridded climatological data instead of global or regional means.
The number of points establishing the relation is dramatically increased from
about 50 points for classical emergent constraints to 237852 points (step 2a)
and 16503 (step 2b) in our approach. We validate the relationship by extensive
out-of-sample testing using a leave-one-model-out \ac{CV} approach and randomly
excluding parts of the data prior to training. Furthermore, the large number of
points in combination with the non-linear \ac{GBRT} model allows us to exploit
more information than classical emergent constraints and include non-linear
multivariate relations in the statistical model, including the information
encoded in spatial variation. The prediction phase of our approach (feeding
observation-based data into the trained \ac{GBRT} model) can be interpreted as
an implicit performance weighting: The \ac{GBRT} model creates predictions
based on today's conditions (in form of variables which are physically
connected to \ac{GPP}). However, unlike other performance weighting techniques
\autocite{Knutti2017a, Sanderson2017}, we do not assign constant weights to the
different climate models. On the contrary, since the algorithm works on grid
cell level, it is able to adapt to regional climate characteristics and thus
applies an implicit \emph{local} and \emph{climate-dependent} weighting.
Moreover, the algorithm adds another level of weighting by assigning a relative
relevance score to the different predictors. We emphasize that this is only an
implicit weighting, since it is not possible to extract specific values for the
individual weights due to the complex structure of the \ac{GBRT} model.

Yet, like emergent constraints and performance weighting techniques, our
\ac{GBRT} approach in step 2 hinges on the assumption that the climate models
reflect the real world. This is certainly only partially true, since there are
processes like \ac{CO2} fertilization, land use and land cover changes,
nutrient cycles and limitations \autocite{Du2020,Fleischer2019}, disturbances,
and induced vegetation dynamics which may strongly alter the future
trajectories of \ac{GPP} without this being fully encoded in today's
observational products. Moreover, all \ac{CO2}-related effects are only
implicitly incorporated in our approach by the target period we are using and
the global rescaling using the emergent constraint in step 1. Thus, our
reduction of uncertainty only applies to the specific setup we are considering
here: the future \ac{GPP} in the \acs{CMIP}5 \acs{RCP}8.5 scenario at the end
of the \nth{21} century. It has to be viewed as a nominal reduction, while the
real uncertainty remains unknown and is likely larger than the nominal one.
These issues could be tested in the future by including other (even offline)
models, which account for the aforementioned processes. A true validation will
only be possible once we experience the changed condition or could be based on
paleo data. In summary, our approach does not address the coupled system
directly but hinges on the climate predictions of the individual \acp{ESM} and
hence is not able to implicitly correct respective biases of the \acs{CMIP}5
ensemble.

The presented approach based on \ac{ML} is not limited to projections of the
future carbon cycle, which is used as an illustrative example. Indeed, since
its only prerequisites are the availability of gridded climate model data and
gridded observation-based data, the \ac{GBRT} algorithm can be applied to any
variable of interest if physically-relevant diagnostics that influence the
target variable are known. This opens a wide range of possibilities for
constraining uncertainties in projected variable scenarios. In addition, maps
of the local relative feature importance as shown in
\cref{fig:06:step2b_ml_inference:c} can be used to reveal connections between
different Earth system variables that are currently unknown. With the proposed
data-driven approach, a possibly extended application could be to constrain
projections of not just one but several variables simultaneously. Moreover, the
concept of our method is not limited to \ac{GBRT} but can be used with any
other (\ac{ML}) regression algorithm.

### 2021-02-22 ch07_summary_outlook/summary_outlook.tex
Initialized chapters
Added first figure
Added figures for ESMValTool contributions (chapter 3)
Added further tables for chapter 05 and text for chapter 06
Finished chapter 06

%=============================================================================%
%                                Dissertation                                 %
%                               Manuel Schlund                                %
%                                  (c) 2020                                   %
%=============================================================================%
%                                  Paper GPP                                  %
%=============================================================================%

\chapter{Constraining Uncertainties in future GPP with Machine Learning}
\label{ch:paper_gpp}

TBA.

>>>>

%=============================================================================%
%                                Dissertation                                 %
%                               Manuel Schlund                                %
%                                  (c) 2020                                   %
%=============================================================================%
%                             Summary and Outlook                             %
%=============================================================================%

\chapter{Summary and Outlook}
\label{ch:07:summary_outlook}

Summary:

- Here we present methods to constrain uncertainties in multi-model projections
of the future climate.

- ESMValTool is great tool to evaluate multi-model output (thus also
projections)

- ESMValTool is expanded and also used to create analyses presented in chapters
4, 5 and 6

- We find high multi-model means and high ranges of ECS and TCR in new CMIP6
ensemble (reasons, specific numbers)

- State-of-the-art methods (emergent constraints) cannot reduce uncertainties
in CMIP6 ECS projections (single-process, more complex models, etc.)

- Add numbers for chapter 5

- Novel ML-based approach: Cannot be used on ECS (scalar values)

- Add numbers for chapter 6

Outlook:

- Why GBRT not used for ECS?

- GBRT used for temperature anomaly based on other studies (emergent
constraints, historical T trend, weighting following Brunner et al.)
